%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% I, the copyright holder of this work, release this work into the
%% public domain. This applies worldwide. In some countries this may
%% not be legally possible; if so: I grant anyone the right to use
%% this work for any purpose, without any conditions, unless such
%% conditions are required by law.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[
  digital,     %% Submit digital version. Change to `printed` for print.
  oneside,     %% One-sided for digital submission. Change to `twoside` for print.
  nosansbold,
  nocolorbold,
  lof,         %% Print List of Figures. Replace with `nolof` to hide.
  lot,         %% Print List of Tables.  Replace with `nolot` to hide.
]{fithesis4}

%% Locale
\usepackage{polyglossia}
\setmainlanguage{english}
\setotherlanguages{czech, slovak}

%% Thesis metadata
\thesissetup{
    date        = \the\year/\the\month/\the\day,
    university  = mu,
    faculty     = fi,
    type        = mgr,
    department  = Department of Software Engineering,
    author      = Bc. Jakub Sikula,
    gender      = m,
    advisor     = {RNDr. Tomáš Rebok, Ph.D.},
    title       = {Design and Implementation of a Platform for Real-Time Environmental Sensor Data Collection, Integration, and Visualisation},
    TeXtitle    = {Design and Implementation of a Platform for Real-Time Environmental Sensor Data Collection, Integration, and Visualisation},
    keywords    = {sensor data, environmental monitoring, OGC SensorThings API, MQTT, TimescaleDB, FastAPI, Next.js, Docker, eLTER},
    TeXkeywords = {sensor data, environmental monitoring, OGC SensorThings~API, MQTT, TimescaleDB, FastAPI, Next.js, Docker, eLTER},
    abstract    = {%
      This thesis designs and implements a platform for real-time environmental sensor data
      collection, integration, and visualisation, targeting the operational needs of Czech
      University of Life Sciences Prague (CZU) within the European eLTER long-term ecosystem
      research infrastructure. The platform builds on the UFZ Time.IO open-source solution,
      which provides MQTT-driven ingestion workers, per-sensor PostgreSQL/TimescaleDB schema
      isolation, and OGC SensorThings API access via FROST-Server. A gap analysis reveals that
      Time.IO lacks project-oriented data management, a user-facing portal, an alert system,
      geospatial layer support, and a functional sensor management service. To address these
      gaps, the thesis contributes three components: (1)~a targeted adaptation of the
      \texttt{tsm-orchestration} stack with local STA views replacing the non-functional
      upstream SMS service, modular Compose structure, and Podman rootless support;
      (2)~\texttt{water-dp}, a new application layer built with FastAPI and Next.js~15 that
      provides project-based sensor management, MQTT-triggered alert evaluation, GeoServer
      integration, and a multilingual web portal; and (3)~\texttt{hydro-deploy}, a unified
      deployment orchestrator merging both stacks via a single configuration file and
      Makefile-based operator interface. The platform is validated through 1,024 automated
      tests and a requirements evaluation covering 24~functional and 8~non-functional
      requirements. The thesis concludes that while the architectural ideas behind
      Time.IO --- per-Thing isolation, MQTT event pipelines, and OGC~STA compliance --- are
      sound, a purpose-built rewrite targeting eLTER-specific requirements is the more
      sustainable path forward than continued adaptation of the upstream codebase.
    },
    declaration = {%
      Hereby I declare that this paper is my original authorial work,
      which I have worked out on my own. All sources, references, and
      literature used or excerpted during elaboration of this work are
      properly cited and listed in complete reference to the due source.

      \medskip

      In accordance with the Masaryk University Directive No.~10/2023 on the
      use of AI tools, I declare that during the preparation of this thesis
      I used AI-based tools --- specifically GitHub Copilot for code
      completion suggestions during development, and ChatGPT for occasional
      consultations on LaTeX formatting and language phrasing. These tools
      were used in a supportive capacity; all design decisions, implementation,
      analysis, and written content are my own work. I have reviewed and take
      full responsibility for the final form and correctness of the entire
      thesis.
    },
    thanks      = {%
      % TODO: Write acknowledgements.
      I would like to thank my advisor, RNDr. Tomáš Rebok, Ph.D., for his guidance throughout this work.
      TODO: complete acknowledgements.
    },
    bib         = thesis.bib,
    facultyLogo = fithesis-fi,
}

%% Packages
\usepackage{makeidx}
\makeindex
\usepackage{paralist}     %% Compact list environments
\usepackage{amsmath}      %% Mathematics
\usepackage{amsfonts}
\usepackage{url}          %% URLs in bibliography
\usepackage{listings}     %% Source code listings
\usepackage{booktabs}     %% \toprule, \midrule, \bottomrule in tables
\usepackage{tabularx}     %% Auto-width table columns
\usepackage{longtable}    %% Tables that span multiple pages
\usepackage{floatrow}     %% Captions above tables
\usepackage[babel]{csquotes} %% Context-sensitive quotation marks
\usepackage{tikz}         %% Diagrams and figures
\usetikzlibrary{shapes, arrows.meta, positioning, fit, backgrounds, calc}
%% microtype is already loaded by fithesis4 with [final,babel,nopatch=item]
\usepackage{pdflscape}     %% Landscape pages in portrait document

\floatsetup[table]{capposition=top}

%% Allow line breaks inside \texttt (monospace) text
\renewcommand{\texttt}[1]{\begingroup\ttfamily\hyphenchar\font=`\-\relax #1\endgroup}
%% Slightly relax line-breaking tolerance to avoid overfull hboxes
\tolerance=1500
\emergencystretch=1.5em

\lstset{
  basicstyle      = \ttfamily\small,
  identifierstyle = \color{black},
  keywordstyle    = \color{blue},
  stringstyle     = \color{teal},
  commentstyle    = \itshape\color{magenta},
  breaklines      = true,
  frame           = single,
  numbers         = left,
  numberstyle     = \tiny\color{gray},
}

\begin{document}

\input{chapters/01-introduction}
\input{chapters/02-background}
\input{chapters/03-requirements}
\input{chapters/04-architecture}
\input{chapters/05-implementation}
\input{chapters/06-testing}
\input{chapters/07-conclusion}

\printbibliography[heading=bibintoc]

\end{document}
