%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% I, the copyright holder of this work, release this work into the
%% public domain. This applies worldwide. In some countries this may
%% not be legally possible; if so: I grant anyone the right to use
%% this work for any purpose, without any conditions, unless such
%% conditions are required by law.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[
  digital,     %% Submit digital version. Change to `printed` for print.
  oneside,     %% One-sided for digital submission. Change to `twoside` for print.
  nosansbold,
  nocolorbold,
  lof,         %% Print List of Figures. Replace with `nolof` to hide.
  lot,         %% Print List of Tables.  Replace with `nolot` to hide.
]{fithesis4}

%% Locale (pdfLaTeX variant — matches Overleaf compiler setting)
\usepackage[resetfonts]{cmap}
\usepackage[T1]{fontenc}
\usepackage[
  main=english,
  english, czech, slovak
]{babel}

%% Thesis metadata
\thesissetup{
    date        = \the\year/\the\month/\the\day,
    university  = mu,
    faculty     = fi,
    type        = mgr,
    department  = Department of Software Engineering,
    author      = Bc. Jakub Sikula,
    gender      = m,
    advisor     = {RNDr. Tomáš Rebok, Ph.D.},
    title       = {Design and Implementation of a Platform for Real-Time Environmental Sensor Data Collection, Integration, and Visualization},
    TeXtitle    = {Design and Implementation of a Platform for Real-Time Environmental Sensor Data Collection, Integration, and Visualization},
    keywords    = {sensor data, environmental monitoring, OGC SensorThings API, MQTT, TimescaleDB, FastAPI, Next.js, Docker, eLTER},
    TeXkeywords = {sensor data, environmental monitoring, OGC SensorThings~API, MQTT, TimescaleDB, FastAPI, Next.js, Docker, eLTER},
    abstract    = {%
      Environmental monitoring networks generate continuous streams of heterogeneous
      sensor data that must be ingested, quality-controlled, stored, and made accessible
      through standard interfaces. This thesis builds on the UFZ Time.IO open-source
      platform to create a working deployment for Czech University of Life Sciences
      Prague within the context of the European eLTER research infrastructure.

      A systematic analysis of Time.IO revealed that key capabilities were missing or
      broken: there was no project-oriented data management, no user-facing portal, and
      the standards-based data access layer was non-functional. The thesis addresses
      these gaps through three components: an adaptation of the upstream ingestion
      stack that restores its core functionality, a new application layer providing
      sensor management, alerting, and a multilingual web portal, and a deployment
      orchestrator that unifies the two stacks into a single operator workflow.

      The result is a containerized platform validated through over a thousand automated
      tests and a full requirements evaluation. The experience shows that the
      architectural ideas behind Time.IO are sound, but a purpose-built implementation
      targeting eLTER's specific needs would be more sustainable than continued
      adaptation of the upstream codebase.
    },
    declaration = {%
      Hereby I declare that this paper is my original authorial work,
      which I have worked out on my own. All sources, references, and
      literature used or excerpted during elaboration of this work are
      properly cited and listed in complete reference to the due source.

      \medskip

      During the preparation of this thesis, the following AI-based tools
      were used in a supportive capacity: GitHub Copilot for code completion
      suggestions during development, Grammarly for grammar and style
      corrections, and Claude for consultations and suggestions. All design
      decisions, implementation, analysis, and written content are my own
      work. I have reviewed and take full responsibility for the final form
      and correctness of the entire thesis.
    },
    thanks      = {%
      I would like to thank my advisor, RNDr. Tom\'{a}\v{s} Rebok, Ph.D., for his
      guidance and feedback throughout this work, and for introducing the
      eLTER context that shaped the platform's direction.

      I also thank prof. Ing. Martin Hanel, Ph.D., from the Faculty of
      Environmental Sciences at the Czech University of Life Sciences Prague,
      for discussions on the hydrological monitoring requirements that
      informed the platform design.
    },
    bib         = thesis.bib,
    facultyLogo = fithesis-fi,
}

%% Packages
\usepackage{makeidx}
\makeindex
\usepackage{paralist}     %% Compact list environments
\usepackage{amsmath}      %% Mathematics
\usepackage{amsfonts}
\usepackage{url}          %% URLs in bibliography
\usepackage{listings}     %% Source code listings
\usepackage{booktabs}     %% \toprule, \midrule, \bottomrule in tables
\usepackage{tabularx}     %% Auto-width table columns
\usepackage{longtable}    %% Tables that span multiple pages
\usepackage{floatrow}     %% Captions above tables
\usepackage[babel]{csquotes} %% Context-sensitive quotation marks
\usepackage{tikz}         %% Diagrams and figures
\usetikzlibrary{shapes, arrows.meta, positioning, fit, backgrounds, calc}
%% microtype is already loaded by fithesis4 with [final,babel,nopatch=item]
\usepackage{pdflscape}     %% Landscape pages in portrait document

\floatsetup[table]{capposition=top}

%% Allow line breaks inside \texttt (monospace) text
\newcommand{\origtt}{}
\let\origtt\texttt
\renewcommand{\texttt}[1]{\origtt{\hyphenchar\font=`\-\relax #1}}
%% Slightly relax line-breaking tolerance to avoid overfull hboxes
\tolerance=1500
\emergencystretch=1.5em

\lstset{
  basicstyle      = \ttfamily\small,
  identifierstyle = \color{black},
  keywordstyle    = \color{blue},
  stringstyle     = \color{teal},
  commentstyle    = \itshape\color{magenta},
  breaklines      = true,
  frame           = single,
  numbers         = left,
  numberstyle     = \tiny\color{gray},
}

\begin{document}

\input{chapters/01-introduction}
\input{chapters/02-background}
\input{chapters/03-requirements}
\input{chapters/04-architecture}
\input{chapters/05-implementation}
\input{chapters/06-testing}
\input{chapters/07-conclusion}

\printbibliography[heading=bibintoc]

\end{document}
