\chapter*{Introduction}
\markright{\textsc{Introduction}}
\addcontentsline{toc}{chapter}{Introduction}
\label{chap:intro}

Environmental monitoring of river systems and catchment areas generates continuous streams of heterogeneous sensor data from
devices produced by different manufacturers, communicating over different protocols, and structured according to different
data models. Managing this diversity at scale --- from data acquisition in the field through ingestion, quality control,
storage, and visualisation --- is a non-trivial engineering problem, particularly for research institutions that operate
many sensors across geographically distributed sites.

Czech University of Life Sciences Prague (CZU) is one such institution. Its sensor infrastructure spans river gauging
stations, catchment monitoring networks, and experimental plots, with instruments recording water level, conductivity,
temperature, turbidity, and other environmental parameters. Prior to this work, no unified platform existed to manage
this infrastructure in a coherent, standards-compliant way.

The UFZ Time.IO solution, developed at the Helmholtz Centre for Environmental Research in Leipzig, was identified as a
candidate foundation. Rather than building a full environmental data management system from scratch, this thesis
investigates whether Time.IO can be adapted to meet the operational requirements of a real institutional deployment,
and implements the extensions needed to close the gap between the upstream system and those requirements.

The platform must also be positioned within the broader context of the European eLTER (Long-Term Ecosystem Research)
infrastructure, which was actively evaluating Time.IO as a candidate for standardised data acquisition across its
network of research sites at the time this work was assigned. The CZU deployment therefore serves a dual purpose:
it addresses a concrete operational need at CZU, and it provides evidence for or against the suitability of the
adapted platform for wider eLTER adoption.

\medskip\noindent
The goals of this thesis are:
\begin{enumerate}
  \item To analyse the capabilities and limitations of the UFZ Time.IO system with respect to the CZU and eLTER
        deployment requirements.
  \item To design and implement the \texttt{water-dp} application layer providing project-oriented sensor management,
        MQTT-driven alert evaluation, geospatial integration, and a multilingual user portal.
  \item To implement the \texttt{hydro-deploy} unified deployment orchestrator enabling single-command installation
        of the complete integrated stack.
  \item To evaluate the resulting platform against the identified requirements and reflect on its suitability for
        broader adoption within the eLTER infrastructure.
\end{enumerate}

\medskip\noindent
The thesis is structured as follows.
Chapter~\ref{chap:background} provides background on environmental sensor data management, the OGC SensorThings API
standard, the UFZ Time.IO solution, and the eLTER infrastructure.
Chapter~\ref{chap:requirements} derives the requirements from the CZU use cases and performs a gap analysis
against the baseline Time.IO system.
Chapter~\ref{chap:architecture} presents the architecture of the resulting platform.
Chapter~\ref{chap:implementation} describes the implementation of all three platform components.
Chapter~\ref{chap:testing} evaluates the platform against the requirements and discusses performance and deployment.
Chapter~\ref{chap:conclusion} summarises the contributions, reflects on limitations, and outlines future work.
