\chapter{Background and Related Work}
\label{chap:background}

Later chapters build on several technologies and prior systems described here.
Section~\ref{sec:bg-sensor-data} reviews environmental sensor data management and its key challenges.
The OGC SensorThings API standard along with its FROST-Server implementation are covered in
Section~\ref{sec:bg-sta}. Section~\ref{sec:bg-timeio} then analyzes the UFZ Time.IO solution
that served as the starting point for the work. The institutional context behind its selection
is discussed in Section~\ref{sec:bg-elter-context}, and Section~\ref{sec:bg-elter} outlines
the eLTER research infrastructure and its data integration requirements.

%% ================================================================
\section{Environmental Sensor Data Management}
\label{sec:bg-sensor-data}
%% ================================================================

Environmental monitoring deployments typically connect sensors to data loggers or gateways
that forward observations to a central infrastructure. A representative architecture has four layers:
edge devices sampling physical quantities, an ingestion layer receiving data over one or more
protocols, a storage layer optimized for time-series workloads, and an access layer exposing the
data through standardized interfaces \parencite{hart2006environmental}.

Heterogeneity across these systems poses a significant engineering challenge. A single monitoring network
may combine sensors from different manufacturers, each using a proprietary payload format, delivered
over MQTT, HTTP~POST, SFTP file upload, or manual CSV import. Temporal resolutions range from
sub-second sampling on vibrating-wire piezometers to daily aggregates from weather stations.
Reconciling these differences into a unified data model is a prerequisite for any cross-site analysis
\parencite{horsburgh2008observations}.

Several platforms address parts of this problem space. The Observations Data Model~2 (ODM2)
\parencite{horsburgh2016odm2} provides a relational schema for managing diverse observational data with
rich provenance metadata. ERDDAP \parencite{simons2023erddap} focuses on data distribution, enabling
scientists to access gridded and tabular datasets through uniform query interfaces. The OGC Sensor
Observation Service (SOS) \parencite{ogcsos} defines a standard web-service interface for querying sensor
data, but its XML-based protocol has been largely superseded by the RESTful SensorThings~API discussed
in Section~\ref{sec:bg-sta}.

What these solutions share is the recognition that interoperability standards are essential for
cross-institutional data sharing. Without a shared data model and access protocol, each
integration turns into a bespoke point-to-point effort. The work in this thesis focuses on
the Czech University of Life Sciences Prague (CZU) deployment, where sensors from multiple
manufacturers need to be unified under a single platform accessible to several research groups.

%% ================================================================
\section{OGC SensorThings API and the FROST Server}
\label{sec:bg-sta}
%% ================================================================

The OGC SensorThings API (STA) \parencite{ogc-sta} is an open standard that defines a RESTful interface
for managing and accessing Internet of Things sensor data. Ratified by the Open Geospatial Consortium in 2016,
it was designed to be lightweight, JSON-based, and suitable for resource-constrained IoT environments ---
in contrast to the earlier XML-heavy Sensor Observation Service.

\subsection{Data Model}
\label{sec:bg-sta-model}

The STA data model is built around six core entity types and their relationships:

\begin{itemize}
  \item \textbf{Thing} --- a physical object of interest, such as a weather station or a water gauge.
        A Thing has a \emph{Location} representing its geospatial position.
  \item \textbf{Datastream} --- a logical grouping of Observations measuring a single
        \emph{ObservedProperty} (e.g.\ water level, air temperature) with a specific \emph{Sensor}.
        Each Datastream belongs to exactly one Thing.
  \item \textbf{Observation} --- an individual measurement event, characterized by a
        \texttt{phenomenonTime} (when the phenomenon occurred), a \texttt{resultTime} (when the
        result was produced), and a \texttt{result} value.
  \item \textbf{Sensor} --- the instrument or procedure that produces Observations.
  \item \textbf{ObservedProperty} --- the physical quantity being measured (e.g.\
        ``Water Temperature'').
  \item \textbf{FeatureOfInterest} --- the real-world feature to which an Observation applies,
        typically derived from the Thing's Location.
\end{itemize}

The standard supports navigation between entities through URL-based association links. For example,
\texttt{/Things(1)/Datastreams} returns all Datastreams belonging to Thing~1, and
\texttt{/Datastreams(5)/Observations?\$filter=phenomenonTime gt 2024-01-01T00:00:00Z}
returns filtered Observations. This query model enables clients to traverse the entity graph
without prior knowledge of the underlying database schema.

STA also defines an MQTT extension (Part~2 of the standard) that allows clients to subscribe to
entity changes in real time, making it suitable for both historical queries and live data streams.

\subsection{FROST-Server}
\label{sec:bg-frost}

As an open-source Java implementation of the OGC SensorThings API,
FROST-Server \parencite{frost-server} was developed by the Fraunhofer Institute for Optronics, System Technologies and Image Exploitation (IOSB).
It is the most widely adopted STA server implementation and serves as the reference for conformance testing.

FROST-Server is deployed as a Java web application on Apache Tomcat and uses PostgreSQL with the
PostGIS extension as its backing store. Each deployment is configured through a Tomcat context file
that specifies the JDBC connection parameters, API version, and plugin settings. The server supports:

\begin{itemize}
  \item Full STA Part~1 (Sensing) compliance, including \texttt{\$filter}, \texttt{\$expand},
        \texttt{\$select}, \texttt{\$orderby}, and \texttt{\$top}/\texttt{\$skip} query options.
  \item The MultiDatastream extension for sensors that produce multiple observed properties
        in a single observation.
  \item Automatic generation of OpenAPI documentation.
  \item Connection pooling with configurable pool sizes (the deployment described in this thesis
        uses a maximum of 20 connections per context).
\end{itemize}

A key architectural property of FROST-Server is that it reads entity data from PostgreSQL views
rather than requiring a specific physical table layout. This allows operators to map an existing
database schema to the STA entity model by defining appropriate SQL views, a capability that
the Time.IO platform exploits through its local STA views, as described in
Section~\ref{sec:bg-timeio-arch}.

%% ================================================================
\section{The UFZ Time.IO Solution}
\label{sec:bg-timeio}
%% ================================================================

Time.IO (internally referred to as \emph{tsm-orchestration} or TSM) is an open-source environmental
sensor data management platform developed at the Helmholtz Centre for Environmental Research (UFZ)
in Leipzig \parencite{bumberger2025timeio,timeio-ref}. It provides an end-to-end pipeline for ingesting, storing,
quality-controlling, and serving environmental time-series data. The platform served as the starting
point for the work described in this thesis.

\subsection{The UFZ TSM Ecosystem}
\label{sec:bg-timeio-ecosystem}

Not a single repository, the UFZ TSM project is a broader ecosystem of over a dozen
components hosted under the \texttt{ufz-tsm} group on the Helmholtz Codebase. Beyond the
core \texttt{tsm-orchestration} ingestion pipeline adopted in this thesis, the ecosystem
includes two frontend projects (\texttt{tsm-frontend}, a Django application with 465~commits,
and \texttt{thing-management}, a newer FastAPI/Vue.js replacement started in September~2024),
a combined deployment repository (\texttt{timeio-and-sms}), the external Sensor Management
System (SMS) discussed in Section~\ref{sec:bg-timeio-limitations}, Kubernetes Helm charts,
developer utilities, and several supporting libraries.

This thesis adopts only \texttt{tsm-orchestration} as its foundation. The upstream frontend
components were replaced by the \texttt{water-dp} portal (Chapter~\ref{chap:implementation})
because both upstream frontends produced invalid MQTT configuration messages that broke the
ingestion pipeline, and neither provided the project-oriented management model required by CZU.

\subsection{Architecture Overview}
\label{sec:bg-timeio-arch}

Time.IO is deployed as a set of Docker containers orchestrated by Docker Compose. The infrastructure
layer comprises the following core services:

\begin{itemize}
  \item \textbf{PostgreSQL with TimescaleDB} --- the central database, storing both the configuration
        database (\texttt{config\_db} schema) and per-thing observation data in isolated schemas.
  \item \textbf{Mosquitto MQTT broker} --- the event backbone connecting all workers and external
        sensor data sources. Supports both plain (port~1883) and TLS-encrypted (port~8883)
        connections.
  \item \textbf{MinIO object storage} --- an S3-compatible store used for file-based data ingestion.
        Each Thing receives its own bucket for raw data files.
  \item \textbf{FROST-Server} --- an OGC SensorThings API endpoint deployed on Apache Tomcat,
        providing standardized read access to observation data through SQL views over per-thing
        schemas.
  \item \textbf{Keycloak} --- an OpenID Connect identity provider managing user authentication
        and role-based access control.
  \item \textbf{Grafana} --- a visualization layer with pre-configured dashboards for time-series
        exploration.
  \item \textbf{Flyway} --- a database migration tool that applies versioned SQL scripts to
        initialise and evolve the database schema.
\end{itemize}

On top of these infrastructure services, Time.IO deploys a set of specialized workers and a
configuration API. The workers follow a message-driven architecture: each subscribes to one or more
MQTT topics, performs its designated processing, and optionally publishes results to downstream
topics. This design decouples the components and allows independent scaling and deployment
of each worker.

Its configuration database (\texttt{config\_db} schema) stores the metadata that drives the entire
pipeline. Key tables include \texttt{thing} (sensor device definitions), \texttt{project} (logical
groupings of Things), \texttt{database} (connection credentials per project), \texttt{s3\_store}
(MinIO bucket configuration), \texttt{file\_parser} and \texttt{file\_parser\_type} (parser
definitions), \texttt{mqtt} and \texttt{mqtt\_device\_type} (MQTT ingestion settings), and
\texttt{ext\_api} and \texttt{ext\_sftp} (external data source configurations).

Central to this design is the \emph{per-thing schema} pattern: when a new Thing is provisioned,
the platform creates a dedicated PostgreSQL schema containing its own observation and
datastream tables. The full provisioning sequence and its isolation properties are described
in Section~\ref{sec:arch-db-perthings}.

Upstream FROST-Server integration relies on a set of SQL views
(\texttt{sql/sta\_views/}) that map the STA entity model to tables managed by the external
Sensor Management System (SMS). These views query SMS-provided metadata (device catalogs,
contact records, deployment actions, and datastream links) to construct the
\texttt{THINGS}, \texttt{DATASTREAMS}, \texttt{OBSERVATIONS}, and related entities that
FROST-Server reads. Because the SMS service was never operational
(Section~\ref{sec:bg-timeio-limitations}), these views returned empty result sets or failed
entirely, rendering the STA endpoint non-functional. Without a working STA layer, FROST-Server
could not serve any observation data, effectively breaking one of the platform's core access
interfaces.

As part of this thesis, an alternative set of local SQL views was introduced to bypass the
SMS dependency. These views derive FROST entities directly from the per-thing schemas and
the configuration database rather than from SMS. Their design is described in
Section~\ref{sec:arch-db-staviews} and their implementation in
Section~\ref{sec:impl-tsm-staviews}.

\subsection{Worker Architecture}
\label{sec:bg-timeio-workers}

All workers inherit from the \texttt{AbstractHandler} base class defined in the \texttt{timeio.mqtt}
module. This base class manages the MQTT connection lifecycle, health-check heartbeats, and
structured error handling. Each worker subscribes to one or more MQTT topics, performs its
designated processing in the \texttt{on\_message} callback, and optionally publishes results
to downstream topics. This message-driven design decouples the components and allows
independent scaling and deployment.

The platform deploys nine workers covering configuration processing, infrastructure
provisioning, data ingestion (file-based and MQTT-based), quality control, and
synchronization with external data sources. A cron scheduler triggers periodic execution
of the synchronization workers. The full worker catalog with MQTT topics and
responsibilities is presented in Table~\ref{tab:tsm-workers}
(Section~\ref{sec:impl-tsm-workers}).

\subsection{Parser Framework}
\label{sec:bg-timeio-parsers}

The data parsing subsystem uses a pluggable architecture based on an abstract base class
(\texttt{AbcParser}) with concrete implementations registered in a parser map. The framework
distinguishes between file parsers (used by the file-ingest worker for uploaded CSV and JSON
files) and MQTT parsers (used by the mqtt-ingest worker for real-time device payloads).

File parsers include a general-purpose CSV parser (\texttt{CsvParser}), a JSON parser
(\texttt{JsonParser}), and a Pandas-based parser (\texttt{PandasParser}) for complex
tabular transformations. MQTT device parsers are specialized for specific hardware:
\texttt{ChirpStackGenericParser} handles LoRaWAN payloads from the ChirpStack network server,
\texttt{CampbellCr6Parser} processes GeoJSON-formatted data from Campbell Scientific CR6 loggers,
\texttt{YdocMl417Parser} handles Ydoc ML-417 water quality sensors, and
\texttt{BrightskyDwdApiParser} parses German Weather Service API responses. A
\texttt{SineDummyParser} generates synthetic sine-wave data for testing.

Each parser produces a list of \texttt{Observation} objects with fields for timestamp, value, origin,
position (identifying the Datastream), and column header. Observations support four result types:
boolean, number, string, and JSON.

Critically, all parsers and device types in the upstream platform are hardcoded in the Python
source code. Adding support for a new sensor model or file format requires implementing a new
parser class, registering it in the parser map, rebuilding the Docker image, and redeploying
the affected worker. There is no mechanism for operators or end users to define new parsers,
MQTT device types, or file-parsing configurations at runtime through a management interface.
The database schema also lacks extensibility: the \texttt{mqtt\_device\_type} and
\texttt{ext\_api\_type} tables have no metadata columns, so any device-specific configuration
must be baked into the parser source code. This rigidity is a significant limitation for
deployments that must accommodate diverse and evolving sensor hardware.

As part of this thesis, several extensions were introduced to address this limitation:
a dynamic parser loader (\texttt{dynamic\_loader.py}) that retrieves user-provided parser
scripts from MinIO at runtime, a \texttt{CustomApiSyncer} class for dynamically loaded
external API synchronization scripts, Flyway migrations adding \texttt{properties} JSON
columns to the \texttt{mqtt\_device\_type} and \texttt{ext\_api\_type} tables
(V2\_27, V2\_29), and a modification to the MQTT ingest worker to pass device-type
properties to the parser at invocation time. These changes are described in
Chapter~\ref{chap:architecture}.

\subsection{Limitations and State at the Time of This Work}
\label{sec:bg-timeio-limitations}

At the time this work commenced, Time.IO exhibited several limitations that directly
motivated the contributions of this thesis. These limitations span the SMS integration, the
upstream frontend solutions, and the overall project maturity.

\paragraph{Sensor Management System.}
The most significant limitation concerned the Sensor Management System (SMS), an external metadata
service intended to provide controlled vocabularies, semantic annotations, and a catalog of sensor
descriptions. The SMS was designed to be the authoritative source of Thing metadata for FROST-Server,
with dedicated SQL views (\texttt{sql/sta\_views/}) mapping SMS-managed entities to the STA data model.
However, the SMS service was never operational. The synchronization worker
(\texttt{sync\_sms\_manager.py}) existed in the codebase, but the upstream SMS backend it depended on
was not available. Without populated SMS tables, the STA views returned empty result sets or failed
to execute, rendering FROST-Server unable to serve any observation data. This was not a minor feature
gap: the STA endpoint is a core access interface of the platform, and its absence made the upstream
TSM deployment effectively non-functional for any client expecting standards-compliant data access.

The local STA views introduced by this thesis (Section~\ref{sec:arch-db-staviews})
restored FROST-Server functionality by deriving entity data from the per-thing schemas and
configuration database rather than from SMS.

The \texttt{timeio-and-sms} combined deployment repository, which was intended to demonstrate
the integrated SMS and Time.IO workflow, proved impossible to deploy successfully. Multiple
dependencies were outdated or no longer available, and the solution contained breaking bugs that
prevented the services from starting. The repository's documentation was partially written in
German and reflected an internal prototype rather than a production-ready integration. With only
175 commits total and activity ceasing in July~2024, the combined SMS deployment did not represent
a viable path forward.

\paragraph{Upstream frontends.}
The original \texttt{tsm-frontend} (Django) provided basic Thing management functionality but
contained bugs that caused inconsistencies between the frontend configuration forms and the
ingestion pipeline's expectations. When a Thing was created or modified through the Django
interface, the resulting MQTT configuration messages did not always match the format and content
that the downstream workers (\texttt{configdb-updater}, \texttt{thing-setup}) expected, leading
to silent failures in infrastructure provisioning.

The newer \texttt{thing-management} replacement (Vue.js~3 + FastAPI) was still in early development
at the time of this work, with limited functionality and similar integration issues. The
frontend's MQTT message generation contained bugs that caused several core ingestion workflows
, particularly file-based ingestion and external API synchronization, to fail. Neither
upstream frontend provided project-based multi-tenancy, role-based access control beyond basic
Helmholtz~AAI group membership, alerting, geospatial layer management, or internationalization
--- all of which were required for the CZU deployment.

\paragraph{Upstream project activity.}
Beyond the specific component issues, the upstream TSM ecosystem showed an overall pattern of
fragmented development. The codebase was distributed across more than a dozen repositories with
varying levels of maturity. The FETA module (\texttt{timeio/feta.py}), which provides an ORM-like
abstraction over the configuration database, was actively maintained and central to the worker
architecture. However, higher-level management features were either absent, buggy, or spread
across multiple partially overlapping projects (\texttt{tsm-frontend} vs.\ \texttt{thing-management}).

The platform lacked several capabilities required for the CZU deployment: project-based
multi-tenancy with role-based access, an alert and notification system, geospatial layer management,
a user-facing portal for sensor configuration and data exploration, and batch CSV import with
configurable parsing. These gaps are analyzed in Chapter~\ref{chap:requirements} and addressed
by the contributions described in Chapters~\ref{chap:architecture} and~\ref{chap:implementation}.

%% ================================================================
\section{Institutional Context: eLTER and the Adoption of Time.IO}
\label{sec:bg-elter-context}
%% ================================================================

Time.IO was not picked after an open competitive evaluation of environmental data management
platforms. Instead, it grew out of active engagement between the European Long-Term Ecosystem Research
(eLTER) leadership and the UFZ development team. At the time, eLTER was looking for candidate platforms
for standardized data acquisition across its growing European network of research sites.

The UFZ, as both a leading Helmholtz research centre and an active participant in eLTER working groups,
proposed Time.IO as a candidate solution. The evaluation framing was practical: could Time.IO serve
real operational deployments beyond UFZ's own sites, and could it satisfy the interoperability
requirements arising from eLTER's cross-site data harmonization goals?

The Czech University of Life Sciences Prague (CZU) deployment described in this thesis serves as a
pilot for this broader assessment. CZU operates hydrological and meteorological monitoring networks
across several catchment areas and needed a platform that could consolidate data from heterogeneous
sensors into a unified, standards-compliant infrastructure. The adoption of Time.IO as the foundation
for this deployment was driven by the eLTER engagement rather than by a systematic comparison of
alternatives. This context is relevant because it explains why the thesis builds on Time.IO despite
its identified limitations: the strategic alignment with eLTER provides a path toward cross-institutional
adoption that would not be available with a platform developed in isolation.

%% ================================================================
\section{eLTER Research Infrastructure}
\label{sec:bg-elter}
%% ================================================================

The European Long-Term Ecosystem, Critical Zone and Socio-Ecological Research Infrastructure
(eLTER~RI) \parencite{elter-ref} is a distributed research infrastructure connecting a network of
environmental monitoring sites across Europe. Its mandate is to enable long-term, cross-site ecosystem
research by providing shared methodologies, data standards, and access protocols.

eLTER encompasses over 400 sites in more than 25 countries, ranging from forest ecosystems and
freshwater catchments to agricultural landscapes and urban environments. The infrastructure provides
two categories of sites: eLTER Sites, which perform long-term observations at the ecosystem level,
and eLTER Platforms, which host intensive measurement campaigns with high spatial and temporal
resolution.

From a data integration perspective, eLTER's requirements are directly relevant to this thesis:

\begin{itemize}
  \item \textbf{Interoperability.} Data from different sites must be queryable through common standards.
        eLTER has identified the OGC SensorThings API as a candidate standard for real-time sensor
        data access, aligning with the STA-based architecture described in this thesis.
  \item \textbf{Metadata harmonization.} Each site must describe its sensors, observed properties,
        and measurement procedures using shared vocabularies. This requirement motivated the SMS
        component discussed in Section~\ref{sec:bg-timeio-limitations}, although its implementation
        remains incomplete.
  \item \textbf{Open data access.} eLTER promotes FAIR data principles (Findable, Accessible,
        Interoperable, Reusable), requiring platforms to expose data through well-documented,
        machine-readable interfaces.
  \item \textbf{Site autonomy.} Individual sites retain operational independence, choosing their
        own sensor hardware and data collection procedures. A platform serving eLTER must therefore
        accommodate heterogeneous hardware configurations rather than mandating a single vendor.
\end{itemize}

The platform described in this thesis addresses these requirements by building on the STA standard
for data access, supporting configurable parsers for diverse sensor hardware, and providing a
deployment model that can be replicated across independent sites. The extent to which the platform
meets eLTER's broader federation and metadata requirements is discussed in
Section~\ref{sec:test-deployment-elter}.
