\chapter{Implementation}
\label{chap:implementation}

This chapter describes the implementation of the three platform components. Section~\ref{sec:impl-tsm} covers the
adaptations made to the \texttt{tsm-orchestration} stack. Section~\ref{sec:impl-backend} describes the
\texttt{water-dp} backend. Section~\ref{sec:impl-frontend} covers the frontend portal. Section~\ref{sec:impl-deploy}
documents the \texttt{hydro-deploy} unified deployment layer.

%% ----------------------------------------------------------------
\section{TSM Orchestration Adaptation}
\label{sec:impl-tsm}
%% ----------------------------------------------------------------

The upstream \texttt{tsm-orchestration} codebase required several targeted adaptations to enable the CZU deployment.
All modifications to upstream files are marked with \texttt{TSM-TEMP-*} identifiers in the codebase to allow clean
identification should a future upstream fix make them redundant. The architectural rationale for these changes
was discussed in Section~\ref{sec:arch-db-staviews}; this section describes their implementation.

\subsection{Local STA Views}
\label{sec:impl-tsm-staviews}

The upstream STA views in \texttt{src/sql/sta\_views/} query tables owned by the external SMS service,
which was non-functional (Section~\ref{sec:bg-timeio-limitations}). The replacement views reside in
\texttt{src/sql/sta\_views\_local/} and query the local \texttt{thing}, \texttt{datastream}, and
\texttt{observation} tables directly.

Seven view files replace the original set. The \texttt{THINGS} view maps the local \texttt{thing} table
to the STA entity columns expected by FROST-Server, appending the Thing's UUID into the JSONB
\texttt{PROPERTIES} column. The \texttt{DATASTREAMS} view extracts unit-of-measurement metadata from the
\texttt{properties} JSONB column and constructs both the flat columns (\texttt{UNIT\_NAME},
\texttt{UNIT\_SYMBOL}, \texttt{UNIT\_DEFINITION}) and the nested JSON object
(\texttt{UNIT\_OF\_MEASUREMENT}) that FROST requires. The \texttt{OBSERVATIONS} view adds a
\texttt{dataSource: local} marker to the output properties to distinguish locally ingested
observations from those originating through SMS. The \texttt{LOCATIONS} view extracts GeoJSON
coordinates from the Thing's \texttt{properties} JSON and constructs both the text-based
\texttt{LOCATION} column and the PostGIS \texttt{GEOM} geometry column.

The remaining entities --- \texttt{SENSORS}, \texttt{FEATURES\_OF\_INTEREST}, and
\texttt{OBSERVED\_PROPERTIES} --- have no local equivalent. Their views return an empty result
set (\texttt{WHERE false}) so that FROST-Server starts without errors but does not expose
fabricated data. A separate \texttt{THINGS\_LOCATIONS} view provides the join table that FROST
uses internally to associate Things with Locations.

A corresponding set of Grafana views in \texttt{src/sql/grafana\_views\_local/} provides
a \texttt{datastream\_properties} view used by the Grafana dashboards to display datastream
metadata without depending on SMS tables.

Every file in the local view set begins with a header comment
\texttt{-{}-~TSM-TEMP-DISABLED}, followed by an instruction to revert by deleting the file
and setting \texttt{USE\_LOCAL\_STA\_VIEWS=false}. At runtime, the thing-setup worker checks
this environment variable (defaulting to \texttt{true}) and deploys the local or upstream
view set accordingly during schema provisioning.

\subsection{Compose Modularisation}
\label{sec:impl-tsm-compose}

To enable \texttt{water-dp} to run alongside TSM without modifying the upstream Compose file, a
dedicated overlay was added:
\texttt{docker-compose-water-dp.yml}. This file defines a single service, \texttt{water-dp-api},
that connects to the TSM database, MQTT broker, FROST-Server, and Keycloak instance. The
\texttt{search\_path} is set to \texttt{water\_dp,public} so that the API can access both its own
tables and the TSM configuration tables. The overlay also starts a Redis container for the Celery
task queue.

The overlay approach ensures that the upstream \texttt{docker-compose.yml} remains unmodified and
can be updated independently. The full Compose file chain used by \texttt{hydro-deploy} is
described in Section~\ref{sec:impl-deploy-compose}.

\subsection{Worker Architecture}
\label{sec:impl-tsm-workers}

All TSM workers extend \texttt{AbstractHandler} from \texttt{timeio/mqtt.py}, which encapsulates
the MQTT connection lifecycle: broker connection, topic subscription, message dispatch, and
self-healthcheck. The healthcheck mechanism publishes a ping to
\texttt{health/<client\_id>/ping} every 60~seconds and monitors a configurable timeout
(default 600~seconds); if no messages are received within the timeout, the worker terminates.

Each worker runs in its own Docker container. There is no central dispatcher; scheduling is
handled by the Mosquitto topic routing and the cron scheduler for periodic tasks. All
\texttt{docker-compose} entrypoints use \texttt{worker\_launcher.sh}, a Bash wrapper that
restarts the Python process up to three times within a configurable window
(\texttt{RESTART\_WINDOW\_SECONDS}, default~10). If the process fails immediately on each
attempt, the launcher exits with status~0 to prevent Docker from triggering an infinite
restart loop.

Table~\ref{tab:tsm-workers} summarises each worker's input topic, output, and responsibility.

\begin{landscape}
{\small
\begin{longtable}{p{0.12\linewidth} p{0.20\linewidth} p{0.13\linewidth} p{0.48\linewidth}}
  \caption{TSM worker processes and their MQTT topics.}
  \label{tab:tsm-workers} \\
  \textbf{Worker} & \textbf{Subscribes To} & \textbf{Publishes} & \textbf{Responsibility} \\
  \hline
  \endfirsthead
  \textbf{Worker} & \textbf{Subscribes To} & \textbf{Publishes} & \textbf{Responsibility} \\
  \hline
  \endhead
  configdb-updater
    & \texttt{frontend\_thing\_update}
    & \texttt{configdb\_update}
    & Validates versioned Thing configuration messages (v4--v7) and upserts them into the
      configuration database. \\[4pt]
  thing-setup
    & \texttt{configdb\_update}
    & ---
    & Orchestrates six provisioning sub-handlers: database schema, MinIO bucket, MQTT
      credentials, Grafana dashboard, FROST instance, and crontab entries. \\[4pt]
  mqtt-ingest
    & \texttt{mqtt\_ingest/\#}
    & \texttt{data\_parsed}
    & Parses raw IoT device messages using the parser framework (Section~\ref{sec:impl-tsm-parsers}),
      resolves the Thing by MQTT username, and inserts observations via the database API. \\[4pt]
  file-ingest
    & \texttt{object\_storage\_notification}
    & \texttt{data\_parsed}
    & Parses uploaded CSV or JSON files from MinIO using \texttt{CsvParser} or
      \texttt{JsonParser} and inserts observations. \\[4pt]
  run-qaqc
    & \texttt{data\_parsed}
    & \texttt{qaqc\_done}
    & Runs configurable quality-control tests against newly parsed observations and writes
      quality flags. \\[4pt]
  sync-extapi
    & \texttt{sync\_ext\_apis}
    & \texttt{data\_parsed}
    & Fetches data from external HTTP APIs (e.g.\ DWD weather service) on a cron schedule. \\[4pt]
  sync-extsftp
    & \texttt{sync\_ext\_sftp}
    & ---
    & Downloads files from external SFTP servers into MinIO, which triggers file-ingest
      via S3 event notification. \\[4pt]
  grafana-user
    & \texttt{userinfo\_keycloak}
    & ---
    & Synchronises Keycloak user accounts with Grafana organisations and roles. \\[4pt]
  monitor-mqtt
    & \texttt{monitor\_mqtt}
    & ---
    & Collects Mosquitto broker metrics (\texttt{\$SYS/broker/}) and writes them to the
      monitoring schema. \\
\end{longtable}
}
\end{landscape}

\subsubsection{Parser Framework}
\label{sec:impl-tsm-parsers}

The parser framework in \texttt{src/timeio/parser/} uses a class hierarchy rooted in
\texttt{AbcParser}, which defines the abstract interface: \texttt{do\_parse()},
\texttt{to\_observations()}, and date accessors.

Two branches extend this base. \texttt{PandasParser} targets file-based ingestion and uses
pandas \texttt{DataFrames} for tabular data. Its subclasses \texttt{CsvParser} and
\texttt{JsonParser} handle CSV files (with configurable timestamp columns and formats) and
JSON files (with \texttt{json\_normalize} support), respectively. \texttt{MqttParser}
targets real-time MQTT messages and returns \texttt{Observation} dataclass instances directly.
Device-specific subclasses handle protocol variations: \texttt{CampbellCr6Parser} for
Campbell~CR6 data loggers, \texttt{BrightskyDwdApiParser} for the DWD weather API,
\texttt{ChirpStackGenericParser} for LoRaWAN ChirpStack payloads, and
\texttt{YdocMl417Parser} for YDOC~ML-417 loggers. A \texttt{SineDummyParser} generates
synthetic sine-wave data for testing.

Parsers are resolved at runtime by the \texttt{get\_parser()} factory function, which looks up
the parser type string in a static registry dictionary. If no match is found, a dynamic loading
fallback fetches a user-provided parser script from MinIO and executes it in an isolated
module namespace.

\subsection{Assessment of the TSM-TEMP Approach}
\label{sec:impl-tsm-assessment}

The \texttt{TSM-TEMP} markers were introduced with the expectation that the upstream SMS
service would eventually become operational and the local views would be replaced. At the
time of writing, the upstream project shows continued development activity but the SMS
service remains non-functional.

From a practical perspective, the local views fulfil the same role as the SMS-backed views
with lower operational complexity: they query tables that the platform already manages, avoid
foreign-table lookups to an external service, and require no additional infrastructure. The
migration path back to upstream views is well-defined --- each affected file documents the
revert procedure in its header comment --- but the diminishing likelihood of an upstream
fix suggests that a targeted rewrite of the affected components is the more defensible
long-term approach.

%% ----------------------------------------------------------------
\section{Water Data Platform --- Backend}
\label{sec:impl-backend}
%% ----------------------------------------------------------------


%   - SQLAlchemy 2.0 async ORM patterns
%   - Alembic migration strategy
%
% Key implementation areas:
%   (a) Project-oriented data model: linking sensors to projects across the TSM schema boundary
%   (b) FROST API client: querying and creating Things/Datastreams without direct DB access
%   (c) Alert evaluation: definitions (threshold, comparison, target datastream),
%                          MQTT-triggered worker, alert state machine
%   (d) GeoServer integration via REST API
%   (f) Keycloak OIDC: token validation middleware, per-project role enforcement

The \texttt{water-dp} API is a FastAPI application that acts as the application-logic layer on top of
\texttt{tsm-orchestration}. It is structured as a layered Python package in \texttt{api/app/}
with four tiers: endpoint routers, Pydantic schemas, service classes, and SQLAlchemy models.

\subsection{Project Structure}
\label{sec:impl-backend-structure}

The \texttt{api/app/} directory follows a conventional layered layout:

\begin{itemize}
  \item \textbf{Endpoints} (\texttt{api/v1/endpoints/}) --- sixteen router modules, each
        responsible for one REST resource. All routers are assembled into a single
        \texttt{APIRouter} in \texttt{api/v1/api.py} and mounted under the
        \texttt{/api/v1} prefix.
  \item \textbf{Schemas} (\texttt{schemas/}) --- Pydantic v2 models for request validation
        and response serialisation. A \texttt{frost/} sub-package provides typed models for
        STA entities (\texttt{Thing}, \texttt{Datastream}, \texttt{Observation}).
  \item \textbf{Services} (\texttt{services/}) --- business-logic classes that the endpoints
        delegate to. A \texttt{timeio/} sub-package encapsulates all integration points with
        the TSM stack: FROST client, MQTT publisher, orchestrator, and direct database access.
  \item \textbf{Models} (\texttt{models/}) --- SQLAlchemy~2.0 ORM classes mapped to the
        \texttt{water\_dp} schema. A shared \texttt{BaseModel} mixin provides audit columns
        (\texttt{created\_at}, \texttt{updated\_at}, \texttt{created\_by},
        \texttt{updated\_by}).
  \item \textbf{Tasks} (\texttt{tasks/}) --- Celery task definitions for asynchronous work:
        alert evaluation, sensor activity recording, inactivity monitoring, and data
        simulation.
  \item \textbf{Worker} (\texttt{worker/}) --- MQTT subscriber that bridges TSM events to
        Celery tasks.
\end{itemize}

The application is created in \texttt{main.py} with a middleware stack applied in order:
\texttt{LoggingMiddleware} (UUID-tagged request logging), \texttt{SecurityHeadersMiddleware}
(security response headers), \texttt{ErrorHandlingMiddleware} (structured JSON error
responses), Starlette \texttt{CORSMiddleware}, \texttt{TrustedHostMiddleware}, and
\texttt{SlowAPIMiddleware} (rate limiting). OpenAPI documentation is exposed only when the
application runs in debug mode.

Four Alembic migration files manage the \texttt{water\_dp} schema evolution: the initial
schema creates tables for projects, dashboards, geospatial layers, alert definitions, alerts,
and jobs; subsequent migrations add the \texttt{sensor\_activity\_configs}
table, the \texttt{project\_members} table for RBAC, and the \texttt{sensor\_id}/\texttt{datastream\_id}
columns on alert definitions. In production, only Alembic manages schema changes; in debug
mode, SQLAlchemy's \texttt{create\_all()} is called at startup for convenience.

\subsection{Project and Sensor Management}
\label{sec:impl-backend-projects}

The central abstraction is the \emph{project}, which groups sensors, alerts, dashboards, and
members into a logical unit. The \texttt{Project} model stores metadata (name, description,
visibility) and references a Keycloak group through \texttt{authorization\_provider\_group\_name}.
Sensors are linked to projects via the \texttt{project\_sensors} association table, which maps
a project UUID to a Thing UUID in the TSM configuration database.

The \texttt{ProjectService} handles project CRUD, member management, and sensor-project
associations. When a sensor is linked to a project, the service resolves the Thing's schema
name through the TSM configuration database, enabling the API to query observations in the
correct per-thing schema. The \texttt{AsyncThingService} provides the core sensor operations
through the asynchronous FROST client: listing Things, fetching Datastreams, and retrieving
time-filtered Observations. A synchronous \texttt{ThingService} variant exists for use in
Celery tasks, where an async event loop is not available.

Sensor creation follows the Thing lifecycle described in Section~\ref{sec:arch-mqtt-lifecycle}.
The \texttt{TimeIOOrchestrator} in \texttt{services/timeio/orchestrator.py} constructs the
configuration payload, derives the schema name from the Keycloak group, generates MQTT and
S3 credentials, and publishes the message to \texttt{frontend\_thing\_update}. The downstream
TSM workers then provision the infrastructure automatically.

\subsection{FROST API Client Integration}
\label{sec:impl-backend-frost}

Two FROST client implementations exist in \texttt{services/timeio/}: a synchronous variant
(\texttt{frost\_client.py}, using \texttt{requests}) and an asynchronous variant
(\texttt{async\_frost\_client.py}, using \texttt{httpx}). Both implement the same method
surface: \texttt{get\_thing()}, \texttt{get\_things()} with STA \texttt{@iot.nextLink}
pagination, \texttt{list\_datastreams()}, \texttt{get\_observations()} with time-range
filtering, and \texttt{get\_locations()}.

The URL scheme follows the FROST multi-tenancy pattern:
\texttt{\{base\_url\}/\{project\_name\}/\{version\}/\{path\}}, where
\texttt{project\_name} maps to a FROST path prefix derived from the Thing's schema.
The synchronous client uses an \texttt{@lru\_cache(maxsize=128)} factory keyed by
\texttt{(base\_url, project\_name, version, frost\_server)} to avoid creating redundant
client instances. The asynchronous client maintains a module-level dictionary cache
with \texttt{httpx.AsyncClient} instances configured for connection pooling (100~maximum
connections, 20~keep-alive).

The asynchronous client is used by \texttt{AsyncThingService} for concurrent FROST
operations, such as fetching Things for multiple projects in parallel during project list
rendering. The synchronous client is used in Celery tasks and in endpoints where the
overhead of an async context switch is not justified.

\subsection{Alert Evaluation System}
\label{sec:impl-backend-alerts}

The alert system consists of definitions, evaluators, and a state machine. An
\texttt{AlertDefinition} is created per project and specifies: the target (a specific
datastream or a QA/QC rule), conditions (field, operator, threshold), and an active flag.

The \texttt{AlertEvaluator} class provides two evaluation entry points:

\begin{enumerate}
  \item \textbf{Sensor data evaluation} --- \texttt{evaluate\_all\_active\_sensor\_rules()}
        periodically queries the latest observation for each targeted datastream and
        evaluates threshold conditions. Conditions specify an operator (\texttt{>},
        \texttt{<}, \texttt{==}) and a threshold value.
  \item \textbf{QA/QC evaluation} --- \texttt{evaluate\_qaqc\_rules\_for\_thing()} queries
        the per-thing observation table directly via \texttt{psycopg2}, computes the ratio
        of flagged observations within a configurable time window, and compares against
        the threshold. Flag levels are classified as \texttt{BAD} (quality value $\geq 255$),
        \texttt{QUESTIONABLE} ($\geq 2$), or \texttt{ANY} ($> 0$).
\end{enumerate}

The MQTT integration (Section~\ref{sec:arch-mqtt-alerts}) bridges TSM events to the
evaluator: the \texttt{mqtt\_handler.py} subscriber receives \texttt{qaqc\_done} messages
and dispatches the \texttt{evaluate\_qaqc\_alerts\_for\_thing} Celery task with the
affected Thing's UUID. Similarly, \texttt{data\_parsed} messages trigger the
\texttt{record\_sensor\_activity} task for inactivity monitoring.

Alert deduplication is enforced by checking for existing active alerts per definition before
creating new ones. QA/QC alerts are automatically resolved when the flagged ratio drops below
the threshold. The three-state lifecycle (\texttt{active} $\to$ \texttt{acknowledged} $\to$
\texttt{resolved}) is described in Section~\ref{sec:arch-mqtt-alerts}.

\subsection{GeoServer Integration}
\label{sec:impl-backend-geoserver}

The \texttt{GeoServerService} in \texttt{services/geoserver\_service.py} wraps the GeoServer
REST~API for programmatic layer management. All operations use HTTP Basic authentication
against the GeoServer admin account.

The service provides methods for workspace creation, PostGIS datastore configuration, and
layer publishing. The \texttt{publish\_layer()} method registers a new feature type with a
specified spatial reference system (SRS), default style, and visibility. The
\texttt{publish\_sql\_view()} method creates layers backed by parameterised SQL views using
GeoServer's \texttt{JDBC\_VIRTUAL\_TABLE} metadata, which enables dynamic per-project
sensor views without pre-creating database views.

Geospatial features are stored in the \texttt{water\_dp} schema using GeoAlchemy2
\texttt{Geometry} columns (SRID~4326). The \texttt{GeoLayer} and \texttt{GeoFeature} models
represent uploaded layers and their individual features. Layer-to-project associations
are managed through the \texttt{LayerProjectAssignment} model.

\subsection{Authentication and Authorisation}
\label{sec:impl-backend-auth}

Token validation is implemented in \texttt{core/security.py}. The module fetches the JWKS
from Keycloak's well-known endpoint and verifies RS256 signatures on incoming JWT tokens.
The JWKS is cached with a one-hour TTL and refreshed under an \texttt{asyncio.Lock} to
prevent concurrent refresh requests. Multi-issuer support allows tokens issued by the
internal Keycloak URL, the Docker-internal address (\texttt{keycloak:8080}), and an optional
external URL to be accepted.

The authentication dependency chain in \texttt{api/deps.py} provides three levels of access
control: \texttt{get\_current\_user()} extracts and verifies the bearer token,
\texttt{has\_role()} checks for a specific Keycloak realm role, and
\texttt{get\_current\_active\_superuser()} verifies administrator privileges.

Per-project authorisation uses the two-tier RBAC model described in
Section~\ref{sec:arch-tech-keycloak}. The \texttt{RBACService} in
\texttt{services/rbac\_service.py} resolves the effective role by evaluating six rules in
priority order: (1)~realm administrator receives owner access to all projects,
(2)~Keycloak group administrator receives owner access to projects in their group,
(3)~an explicit \texttt{project\_members} row grants the stored role,
(4)~the original project creator is treated as owner for backward compatibility,
(5)~any member of the project's Keycloak group is granted viewer access, and
(6)~if none of the above match, the request is rejected with HTTP~403. The JWT \texttt{groups} claim is parsed by
\texttt{parse\_group\_roles()}, which extracts the Keycloak subgroup suffix
(e.g.\ \texttt{/GroupName/editors} yields role \texttt{editor}). The resolved role determines
the capability set: \texttt{can\_view}, \texttt{can\_edit\_settings},
\texttt{can\_edit\_alerts}, \texttt{can\_link\_sensors}, \texttt{can\_manage\_members},
\texttt{can\_delete}, \texttt{global\_sms\_access}, and \texttt{global\_layers\_access}.

%% ----------------------------------------------------------------
\section{Water Data Platform --- Frontend Portal}
\label{sec:impl-frontend}
%% ----------------------------------------------------------------

The frontend portal is a Next.js~15 application using the App~Router with React~19. The
application is deployed under the \texttt{/portal} base path and built in \texttt{standalone}
output mode for containerised deployment.

\subsection{Information Architecture}
\label{sec:impl-frontend-ia}

The routing structure follows the App~Router convention, where each directory under
\texttt{app/} corresponds to a URL segment. The major route groups are:

\begin{itemize}
  \item \textbf{\texttt{/projects}} --- project list, creation, and per-project views. Each
        project page (\texttt{/projects/[id]}) provides sub-routes for sensor data
        (\texttt{/data}), individual sensor time-series (\texttt{/data/[sensorId]}),
        dashboards (\texttt{/dashboards}), alerts, QA/QC configuration, sensor simulator,
        and project settings.
  \item \textbf{\texttt{/sms}} --- Sensor Management System interface for sensor, parser,
        device type, and external API type administration. Access is restricted to users
        with at least the \texttt{editor} role.
  \item \textbf{\texttt{/layers}} --- GeoServer layer management: listing, upload, and
        detail views.
  \item \textbf{\texttt{/settings}} --- user profile and password management.
\end{itemize}

All authenticated routes use a protected layout pattern. Each layout calls the server-side
\texttt{auth()} function from NextAuth and redirects unauthenticated users to the sign-in
page. The SMS and Layers layouts additionally check the JWT group role to enforce the minimum
\texttt{editor} privilege. The root layout wraps the entire application in a \texttt{Providers}
component that supplies the React~Query \texttt{QueryClientProvider}, the NextAuth
\texttt{SessionProvider}, a theme context, and an idle-timeout monitor.

\subsection{Time-Series Visualisation}
\label{sec:impl-frontend-charts}

The \texttt{TimeSeriesChart} component renders multi-series time-series data using the
Recharts library. The component supports interactive zoom via two mechanisms: a
\texttt{ReferenceArea} drawn by click-and-drag to select a time range, and a
\texttt{Brush} component for scrolling through the full dataset. Multiple series
(each with its own colour, label, and unit) are merged into a unified time axis by indexing
all data points into a shared timestamp map.

Data is fetched through React~Query hooks in \texttt{hooks/queries/}, which provide
caching, background refetching, and loading-state management. Ten query-hook modules
cover the major resource types: projects, sensors, dashboards, alerts, QA/QC,
SMS, layers, simulator, and permissions. Query keys are centralised in
\texttt{hooks/queries/keys.ts} to ensure cache consistency across components.

The API client (\texttt{lib/api.ts}) is an Axios instance with a request interceptor that
attaches the NextAuth access token as a \texttt{Bearer} header. An in-memory token cache
with a 60-second expiry buffer avoids redundant \texttt{getSession()} calls.

\subsection{Geospatial Layer Viewer}
\label{sec:impl-frontend-geo}

Geospatial rendering uses MapLibre~GL~JS in three components. The \texttt{ProjectMap}
component displays sensor locations as clustered GeoJSON markers on a CARTO basemap
(dark-matter or positron, selected by the theme context). Marker popups show the latest
datastream values, refreshed every 30~seconds. GeoServer WMS layers assigned to the
project are overlaid as raster tiles, and a layer-switching menu allows users to
toggle their visibility.

The \texttt{LocationPickerMap} component provides an interactive click-or-drag marker for
setting sensor coordinates during Thing creation. The \texttt{SensorMiniMap} component
renders a read-only map in sensor detail views showing the sensor's location.

\subsection{Data Quality Interface}
\label{sec:impl-frontend-quality}

The QA/QC interface is accessible at two levels: per-project (\texttt{/projects/[id]/qaqc})
and globally through the SMS section (\texttt{/sms/qa-qc}). The \texttt{SensorQAQCSection}
component renders the QC test configuration for a single sensor, allowing users to define
test types, thresholds, and parameters. A dedicated API client (\texttt{lib/qaqc-api.ts})
provides typed functions for three endpoint scopes: SMS-level, Thing-level, and
project-level QA/QC configuration.

\subsection{Batch Import}
\label{sec:impl-frontend-batch}

The \texttt{BulkImportDialog} component provides a drag-and-drop CSV upload workflow for
creating multiple sensors in a single operation. The \texttt{DataUploadModal} and
\texttt{DatasetUploadModal} components handle historical observation data import: users
upload CSV files which are stored in the per-Thing MinIO bucket, triggering the TSM
file-ingest pipeline. The per-sensor data view also includes a \texttt{FileUploadDialog}
for individual file uploads.

\subsection{Multilingual Support}
\label{sec:impl-frontend-i18n}

The portal supports three languages: English, Czech, and Slovak. The implementation uses
a custom solution based on a Zustand store (\texttt{useLanguageStore}) with
\texttt{persist} middleware, which saves the selected language to \texttt{localStorage}
under the key \texttt{language-storage}. Locale files in \texttt{locales/} export typed
\texttt{Dictionary} objects (\texttt{en.ts}, \texttt{cs.ts}, \texttt{sk.ts}) that share
a common interface definition.

The \texttt{useTranslation()} hook returns a \texttt{t(path, params?)} function that
traverses the dictionary by dot-notation keys (e.g.\ \texttt{t('header.projects')}) and
supports \texttt{\{\{param\}\}} interpolation. The \texttt{LanguageSwitcher} component in
the application header allows users to switch the language at runtime.

Locale selection is purely client-side; there is no server-side locale detection or
URL-based language routing. This design was chosen to simplify the deployment and avoid
additional Next.js middleware complexity, at the cost of the initial page flash before
the persisted language is loaded from storage.

%% ----------------------------------------------------------------
\section{Hydro Deploy --- Unified Deployment Layer}
\label{sec:impl-deploy}
%% ----------------------------------------------------------------

The \texttt{hydro-deploy} repository provides a single entry point for deploying the full integrated platform.
It does not introduce any new runtime services; its sole responsibility is assembly and configuration.
The architectural design --- Compose file merging, secret management, Makefile targets, and
Podman support --- was described in Section~\ref{sec:arch-deployment}. This section covers
implementation details not addressed there.

\subsection{Compose Multi-File Merge}
\label{sec:impl-deploy-compose}

The Compose file chain is described architecturally in Section~\ref{sec:arch-deploy-compose}.
In the Docker variant, four files are merged:

\begin{enumerate}
  \item \texttt{tsm-orchestration/docker-compose.yml} --- all TSM core services.
  \item \texttt{water-dp/docker-compose.yml} --- standalone water-dp services.
  \item \texttt{water-dp/docker-compose.tsm.yml} --- overlay that redirects water-dp
        services to TSM's database, MQTT broker, and Keycloak.
  \item \texttt{hydro-deploy/docker-compose.yml} --- the final overlay, which maps all
        three network names (\texttt{default}, \texttt{water\_shared\_net},
        \texttt{tsm\_network}) to the single external network \texttt{hydro-platform-net},
        injects environment variables, and uses the YAML \texttt{!override} tag to replace
        development bind-mount volumes with production-ready empty lists.
\end{enumerate}

In the Podman variant, a parallel set of \texttt{.podman.yml} files generated by
\texttt{podman\_prep.py} (Section~\ref{sec:arch-deploy-podman}) is used instead.

The startup is staged by \texttt{scripts/start-docker.sh} (or \texttt{start-podman.sh}),
which brings services up in dependency order: TSM core first (waiting for the database to
become healthy), then \texttt{postgres-app}, then GeoServer (with a 300-second timeout),
then the API, and finally the worker, frontend, and optional Cloudflare tunnel.

\subsection{Secrets and Environment Configuration}
\label{sec:impl-deploy-env}

A single \texttt{.env} file is shared across all three repositories via symbolic links
created during setup. The \texttt{scripts/generate-secrets.sh} script populates
cryptographic credentials using Python's \texttt{secrets} module. Three generator functions
are used: \texttt{hex32()} for 32-byte hex tokens, \texttt{fernet\_key()} for Fernet
encryption keys (with a fallback to \texttt{base64}-encoded random bytes if the
\texttt{cryptography} library is unavailable), and \texttt{passphrase()} for 24-character
alphanumeric strings.

Credentials that must be identical across stacks are managed in synchronisation groups: a
single generated database password is written to \texttt{DATABASE\_PASSWORD},
\texttt{TIMEIO\_DB\_PASSWORD}, \texttt{MQTT\_AUTH\_POSTGRES\_PASS}, and several Flyway
variables simultaneously. The script is safe to re-run: it replaces only values that still
contain the \texttt{changeme} placeholder.

Environment consistency is verified by the \texttt{scripts/check.sh} health-check script.
Five \texttt{@sync} variable pairs are validated for equality (e.g.\
\texttt{DATABASE\_PASSWORD} and \texttt{TIMEIO\_DB\_PASSWORD}). The script also checks all
containers for health status, probes HTTP endpoints for reachability, and warns about
remaining \texttt{changeme} placeholders. The derived variables (\texttt{PROXY\_URL},
\texttt{KEYCLOAK\_URL}, etc.) are computed from \texttt{PUBLIC\_HOSTNAME} and
\texttt{PUBLIC\_PORT} by the \texttt{make update-env} target using \texttt{sed}
substitutions.
