\chapter{Testing and Evaluation}
\label{chap:testing}

The platform is evaluated against the requirements from Chapter~\ref{chap:requirements}.
Section~\ref{sec:test-approach} outlines the testing strategy used. The requirements
are assessed one by one in Section~\ref{sec:test-requirements}. Performance and scalability
considerations follow in Section~\ref{sec:test-performance}, and Section~\ref{sec:test-deployment}
covers the CZU deployment and the path to eLTER adoption.

%% ----------------------------------------------------------------
\section{Testing Approach}
\label{sec:test-approach}
%% ----------------------------------------------------------------

The platform is tested at three levels: automated unit and integration tests for the
\texttt{water-dp} backend and frontend, automated tests inherited from the upstream
\texttt{tsm-orchestration} codebase, and manual end-to-end verification of the full
deployment.

\subsection{Backend Test Suite}
\label{sec:test-approach-backend}

The \texttt{water-dp} backend includes a pytest-based test suite comprising 797~test
functions across 43~test files, organized into five directories:
\texttt{test\_api/} (endpoint-level tests), \texttt{test\_services/} (business-logic tests),
\texttt{test\_core/} (middleware, security, and exception handling), \texttt{test\_schemas/}
(Pydantic model validation), and \texttt{test\_tasks/} (Celery background tasks).

The test fixtures are defined in \texttt{conftest.py}. A \texttt{mock\_db\_session} fixture
provides a \texttt{MagicMock(spec=Session)} for SQLAlchemy interactions. The \texttt{client}
fixture creates a FastAPI \texttt{TestClient} with dependency overrides that inject the mocked
database session and an authenticated admin user, enabling endpoint tests to run without a
real database or Keycloak instance. An autouse \texttt{disable\_seeding} fixture prevents
the application from running database initialization at startup.

Custom pytest markers (\texttt{api}, \texttt{services}, \texttt{core}, \texttt{v1}) are
assigned automatically based on the test file's directory, enabling selective test execution.
The default configuration in \texttt{pytest.ini} excludes tests marked as
\texttt{integration}, which require a live PostgreSQL instance.

Coverage is measured with \texttt{pytest-cov} against the \texttt{app/} package. The
largest test files target the most critical business logic: the TimeIO database service
(152~tests covering schema management and observation queries), the RBAC permission resolver
(76~tests covering role inheritance, batch resolution, and group-based authorization),
and the alert evaluator (69~tests covering all operators, deduplication, sensor data
evaluation, and QA/QC rule evaluation).

A Locust-based load testing suite is included in \texttt{tests/load/locustfile.py}. It
simulates authenticated users performing weighted API operations (project listing, sensor
queries, observation retrieval) against a running instance. The load test can be scaled
via \texttt{docker-compose.loadtest.yml}, which supports multiple Locust worker containers.

\subsection{Frontend Test Suite}
\label{sec:test-approach-frontend}

The frontend includes 96~test cases in 13~files under \texttt{frontend/\_\_tests\_\_/},
executed with Vitest and React Testing Library against a jsdom environment. The tests cover
the API interceptor (authentication header injection and error handling), React~Query hook
factories (query key generation, mutation lifecycle), and selected component rendering
(project cards). The Vitest configuration uses the \texttt{@vitejs/plugin-react} plugin
and resolves import aliases through the same path mapping used in production.

\subsection{TSM Orchestration Tests}
\label{sec:test-approach-tsm}

The upstream \texttt{tsm-orchestration} repository contains 131~test functions covering
parser implementations (CSV, JSON), worker scripts (file ingest, crontab setup), external
API syncers (DWD, TTN, Bosch, T-Systems, UBA), Grafana API integration, and cryptographic
utilities. These tests were inherited with the fork and continue to pass against the
modified codebase.

\subsection{Manual End-to-End Verification}
\label{sec:test-approach-e2e}

Full-stack scenarios are verified manually against a running deployment. The primary
verification path traces the complete Thing lifecycle: sensor creation through the portal,
infrastructure provisioning by TSM workers (observable through container logs and database
schema creation), data ingestion via MQTT or CSV upload, observation availability through
the FROST-Server STA endpoint, and time-series rendering in the portal charts. The
\texttt{make check} target in \texttt{hydro-deploy} automates the infrastructure-level
portion of this verification by probing container health, HTTP endpoint reachability, and
\texttt{@sync} variable consistency.

%% ----------------------------------------------------------------
\section{Evaluation Against CZU Requirements}
\label{sec:test-requirements}
%% ----------------------------------------------------------------

Each functional and non-functional requirement from Chapter~\ref{chap:requirements} is
evaluated below. Requirements are classified as \emph{fully met}, \emph{partially met},
or \emph{not met}, with evidence from the test suite, source code, or deployed system.

\subsection{Functional Requirements}
\label{sec:test-requirements-functional}

Table~\ref{tab:req-evaluation-fr} summarizes the evaluation of functional requirements.

\begin{landscape}
{\small
\begin{longtable}{l p{0.30\linewidth} l p{0.38\linewidth}}
  \caption{Functional requirements evaluation.}
  \label{tab:req-evaluation-fr} \\
  \textbf{ID} & \textbf{Requirement} & \textbf{Status} & \textbf{Evidence} \\
  \hline
  \endfirsthead
  \textbf{ID} & \textbf{Requirement} & \textbf{Status} & \textbf{Evidence} \\
  \hline
  \endhead
  FR-1  & MQTT real-time ingestion with per-Thing parsers
        & Fully met
        & mqtt-ingest worker; parser framework with 5~device-specific parsers;
          131~upstream tests pass. \\[3pt]
  FR-2  & CSV file upload ingestion
        & Fully met
        & file-ingest worker; 13~\texttt{test\_things.py} tests cover CSV upload
          endpoint. \\[3pt]
  FR-3  & External SFTP synchronization
        & Fully met
        & sync-extsftp worker; triggered by cron via MQTT. \\[3pt]
  FR-4  & External REST API synchronization
        & Fully met
        & sync-extapi worker; DWD, TTN, Bosch, T-Systems, UBA syncers;
          30~upstream syncer tests. \\[3pt]
  FR-5  & Custom parser and syncer script upload
        & Fully met
        & Dynamic loader fetches scripts from MinIO; AST-based validation blocks
          dangerous operations. \\[3pt]
  FR-6  & Sensor CRUD with parser and coordinate config
        & Fully met
        & 25~\texttt{test\_sms.py} tests; 31~\texttt{test\_orchestrator.py}
          tests. \\[3pt]
  FR-7  & Bulk sensor import from CSV template
        & Fully met
        & \texttt{BulkImportDialog} component; 6~\texttt{test\_bulk.py} tests
          with validation and authorization checks. \\[3pt]
  FR-8  & Sensor inactivity detection and alerting
        & Fully met
        & \texttt{MonitoringService}; 51~\texttt{test\_monitoring\_service.py}
          tests. \\[3pt]
  FR-9  & Reusable MQTT device types and CSV parsers
        & Fully met
        & SMS catalog endpoints; parser/device-type CRUD tested in
          \texttt{test\_sms.py}. \\[3pt]
  FR-10 & Project CRUD with sensor grouping
        & Fully met
        & 10~\texttt{test\_project\_service.py} tests;
          6~\texttt{test\_projects.py} endpoint tests. \\[3pt]
  FR-11 & Project member roles (owner/editor/viewer)
        & Fully met
        & Two-tier RBAC; 76~\texttt{test\_rbac\_service.py} tests cover role
          inheritance and batch resolution. \\[3pt]
  FR-12 & Per-project TimeIO schema provisioning
        & Fully met
        & \texttt{TimeIOOrchestrator}; 152~\texttt{test\_timeio\_db.py} tests
          cover schema management. \\[3pt]
  FR-13 & QA/QC test pipeline configuration
        & Fully met
        & QA/QC endpoints at project and SMS scope;
          \texttt{SensorQAQCSection} frontend component. \\[3pt]
  FR-14 & Automatic QA/QC on \texttt{data\_parsed} event
        & Fully met
        & run-qaqc worker subscribes to \texttt{data\_parsed} topic. \\[3pt]
  FR-15 & On-demand QA/QC triggering via API
        & Fully met
        & QA/QC service publishes to \texttt{data\_parsed} to trigger
          re-evaluation. \\[3pt]
  FR-16 & Interactive time-series charts
        & Fully met
        & \texttt{TimeSeriesChart} component with Recharts; zoom via
          \texttt{ReferenceArea} and \texttt{Brush}. \\[3pt]
  FR-17 & Interactive map with sensor markers and WMS overlay
        & Fully met
        & \texttt{ProjectMap} with MapLibre~GL; GeoServer WMS layer overlay;
          30-second auto-refresh. \\[3pt]
  FR-19 & CSV export of Observation data
        & Fully met
        & Export endpoint in \texttt{things.py} router. \\[3pt]
  FR-20 & OIDC authentication via Keycloak
        & Fully met
        & JWKS validation in \texttt{core/security.py}; 4~\texttt{test\_auth.py}
          tests; 7~\texttt{test\_keycloak\_service.py} tests. \\[3pt]
  FR-21 & Two-tier RBAC model
        & Fully met
        & \texttt{RBACService} with six-rule priority chain;
          76~permission tests. \\[3pt]
  FR-22 & Keycloak group management
        & Partially met
        & Group management decommissioned from portal; users manage groups
          via the Keycloak admin console directly. \\[3pt]
  FR-23 & Historical CSV data upload (up to 200~MB)
        & Fully met
        & \texttt{DataUploadModal} and \texttt{DatasetUploadModal} components;
          files stored in per-Thing MinIO bucket. \\[3pt]
  FR-24 & GeoJSON geospatial import (up to 200~MB)
        & Fully met
        & Bulk import endpoint; 6~\texttt{test\_bulk.py} tests cover streaming
          and validation. \\[3pt]
  FR-25 & Multilingual UI (EN, CS, SK)
        & Fully met
        & Zustand-persisted locale store; typed dictionary files for all three
          languages. \\
\end{longtable}
}
\end{landscape}

\begin{landscape}
\subsection{Non-Functional Requirements}
\label{sec:test-requirements-nonfunctional}

Table~\ref{tab:req-evaluation-nfr} summarizes the evaluation of non-functional requirements.

{\small
\begin{longtable}{l p{0.30\linewidth} l p{0.38\linewidth}}
  \caption{Non-functional requirements evaluation.}
  \label{tab:req-evaluation-nfr} \\
  \textbf{ID} & \textbf{Requirement} & \textbf{Status} & \textbf{Evidence} \\
  \hline
  \endfirsthead
  \textbf{ID} & \textbf{Requirement} & \textbf{Status} & \textbf{Evidence} \\
  \hline
  \endhead
  NFR-1 & Containerized deployment (Docker + Podman)
        & Fully met
        & \texttt{hydro-deploy} supports both runtimes; \texttt{podman\_prep.py}
          applies seven transformations. \\[3pt]
  NFR-2 & OGC STA compliance
        & Partially met
        & FROST-Server serves STA queries; local STA views provide entity data
          but stub views for \texttt{SENSORS}, \texttt{OBS\_PROPERTIES}, and
          \texttt{FEATURES\_OF\_INTEREST} return empty results. \\[3pt]
  NFR-3 & OIDC standards compliance
        & Fully met
        & RS256 JWKS validation with one-hour TTL cache; multi-issuer support;
          \texttt{asyncio.Lock} for concurrent refresh. \\[3pt]
  NFR-4 & Modular extensibility
        & Fully met
        & Dynamic parser loading from MinIO; runtime-uploadable syncer
          scripts. \\[3pt]
  NFR-5 & Security (script sandboxing, rate limiting, encryption)
        & Fully met
        & AST-based script validation; SlowAPI rate limiting; Fernet credential
          encryption; 10~\texttt{test\_crypto\_utils.py} tests. \\[3pt]
  NFR-6 & Maintainability (migrations, OpenAPI docs)
        & Fully met
        & Four Alembic migrations for \texttt{water\_dp} schema; Flyway for
          TSM; OpenAPI docs generated automatically by FastAPI. \\[3pt]
  NFR-7 & eLTER interoperability alignment
        & Partially met
        & STA-based data access enables standard queries; per-project isolation
          supports multi-site operation; eLTER-specific metadata schemas not
          yet implemented. \\[3pt]
  NFR-8 & Unified deployment via single config
        & Fully met
        & Single \texttt{.env} with symlinks; \texttt{make up} deploys the
          full stack; \texttt{make check} validates health. \\
\end{longtable}
}
\end{landscape}

\subsection{Gap Analysis Closure}
\label{sec:test-requirements-gaps}

Table~\ref{tab:gap-closure} maps the eleven gaps identified in
Section~\ref{sec:req-gap} to their resolution status.

\begin{longtable}{l l p{0.55\textwidth}}
  \caption{Gap analysis closure.}
  \label{tab:gap-closure} \\
  \textbf{Gap} & \textbf{Status} & \textbf{Resolution} \\
  \hline
  \endfirsthead
  \textbf{Gap} & \textbf{Status} & \textbf{Resolution} \\
  \hline
  \endhead
  G-1  & Resolved & Project model with member roles and sensor assignment
         implemented in the \texttt{water\_dp} schema. \\[3pt]
  G-2  & Resolved & Next.js portal with time-series charts, maps, alerts,
         QA/QC, and sensor management. \\[3pt]
  G-3  & Resolved & GeoServer integration with REST~API layer management and
         MapLibre~GL WMS overlay. \\[3pt]
  G-4  & Resolved & Nine local STA views replace the SMS-dependent views;
         FROST-Server starts and serves valid STA responses. \\[3pt]
  G-5  & Resolved & \texttt{water-dp} replaces the non-functional SMS with
         sensor management endpoints and the \texttt{TimeIOOrchestrator}. \\[3pt]
  G-6  & Resolved & Alert evaluation system with sensor threshold, QA/QC, and
         inactivity rules; MQTT-triggered processing. \\[3pt]
  G-7  & Resolved & \texttt{hydro-deploy} with Compose multi-file merge,
         \texttt{@sync} validation, and Makefile interface. \\[3pt]
  G-8  & Resolved & \texttt{podman\_prep.py} generates Podman-compatible Compose
         files; rootless-specific UID/GID and network fixes applied. \\[3pt]
  G-9  & Resolved & Three locale files (EN, CS, SK) with Zustand-persisted
         language switching. \\[3pt]
  G-11 & Resolved & \texttt{TimeIOOrchestrator} generates MQTT configuration
         messages compatible with the ingestion pipeline. \\[3pt]
  G-12 & Resolved & SMS catalog endpoints for parser and device-type CRUD;
         dynamic parser loading from MinIO. \\
\end{longtable}

All eleven gaps from the baseline analysis are resolved. Two non-functional requirements
(NFR-2 and NFR-7) remain partially met: NFR-2 due to the stub STA entity views that
return empty results for entities without a local equivalent, and NFR-7 due to the
absence of eLTER-specific metadata schemas, which are planned as future work.

%% ----------------------------------------------------------------
\section{Performance and Scalability Considerations}
\label{sec:test-performance}
%% ----------------------------------------------------------------

No formal performance benchmarking was carried out as part of this thesis.
What follows is a discussion of architectural properties that affect performance, together
with the known bottlenecks.

\subsection{Time-Series Query Performance}
\label{sec:test-perf-timescaledb}

TimescaleDB partitions each per-thing observation table into time-based chunks. Time-range
queries --- the dominant access pattern for chart rendering --- benefit from chunk exclusion:
the query planner skips chunks outside the requested interval, reducing I/O to the relevant
time window. Older chunks are compressed automatically, further reducing storage requirements
and improving sequential scan throughput for historical queries.

The per-thing schema isolation provides an additional performance benefit: queries for one
sensor's data never scan another sensor's observation table, even when many sensors are
onboarded. This eliminates cross-sensor interference regardless of data volume.

\subsection{API Throughput}
\label{sec:test-perf-api}

The FastAPI backend uses asynchronous request handlers with \texttt{httpx}-based FROST
client calls, enabling the API to serve multiple concurrent requests without blocking on
FROST-Server responses. The \texttt{AsyncFrostClient} maintains a connection pool
(100~maximum connections, 20~keep-alive) to reduce TCP handshake overhead for repeated
FROST queries.

The Locust load testing suite included in the repository enables throughput measurement
under simulated multi-user load. The tests authenticate against Keycloak and exercise the
primary API operations (project listing, sensor queries, observation retrieval) with
configurable concurrency.

\subsection{Known Bottleneck}
\label{sec:test-perf-bottleneck}

FROST-Server represents the main performance bottleneck. As a synchronous Java application
running on Tomcat, it processes STA queries sequentially within each thread. When the
\texttt{water-dp} API fetches sensor metadata for a project with many Things, each
FROST query must finish before the next one starts. \texttt{AsyncThingService}
mitigates this by firing FROST requests concurrently via \texttt{httpx} connection pooling,
but FROST-Server itself remains the throughput ceiling. Swapping FROST for a lighter STA
implementation or adding a caching layer for frequently accessed metadata are potential
future optimizations.

%% ----------------------------------------------------------------
\section{Deployment on CZU Infrastructure and eLTER Path}
\label{sec:test-deployment}
%% ----------------------------------------------------------------

The platform was deployed on CZU infrastructure as the primary validation environment
for this thesis. This section documents the deployment, evaluates the platform's
readiness for adoption within the eLTER research infrastructure, and assesses the
long-term sustainability of the TSM-TEMP workarounds.

\subsection{CZU Deployment}
\label{sec:test-deployment-czu}

The deployment workflow described in Sections~\ref{sec:arch-deployment}
and~\ref{sec:impl-deploy} was validated on CZU infrastructure. The full stack of
16~containers was started using \texttt{make up}, with the post-deployment
\texttt{check.sh} script confirming container health, endpoint reachability across
all eight HTTP services, and credential consistency across the synchronised
\texttt{.env} variable pairs.

A platform audit conducted during this work identified and remediated 14~critical and
high-severity issues across the water-dp codebase, including security header
enforcement, CORS and debug-mode controls, and credential management. An additional
27~issues were documented in the upstream TSM codebase, where remediation was not
feasible without upstream cooperation. The audit assessed all ten OWASP~Top~10
categories; accepted deferrals include the absence of a circuit-breaker pattern and
MQTT quality-of-service limited to level~1 by the upstream broker configuration.

\subsection{Path to eLTER Adoption}
\label{sec:test-deployment-elter}

Several architectural properties of the platform align with eLTER requirements
described in Section~\ref{sec:bg-elter}. The data access layer is built on the OGC
SensorThings API, which is a candidate standard for eLTER data exchange. Authentication
uses Keycloak with OpenID Connect, supporting the federated identity models common in
European research infrastructures. Project-based isolation provides the multi-tenancy
needed for hosting data from independent research sites. The containerized deployment
with single-file configuration enables reproducible installation across heterogeneous
institutional environments.

However, no eLTER-specific features have been implemented. The following gaps remain
between the current platform and a production-ready eLTER deployment:

\begin{itemize}
  \item \textbf{Metadata schemas.} eLTER defines site-level metadata (DEIMS-SDR
    identifiers, habitat classifications, geographic delineations) that have no
    representation in the current data model. Ingesting eLTER metadata would require
    extending the project and sensor models or introducing a dedicated metadata service.
  \item \textbf{Cross-site federation.} The platform operates as a single-site
    deployment. eLTER's distributed nature, spanning over 400~sites across more than
    25~countries, requires federation mechanisms: cross-site discovery, aggregated
    queries, and data harmonization across independently operated instances.
  \item \textbf{Vocabulary alignment.} eLTER uses controlled vocabularies (EnvThes)
    for observed properties and ecosystem parameters. The current platform stores
    observed-property metadata as free-text fields inherited from FROST, with no
    vocabulary binding or validation.
  \item \textbf{Provenance and FAIR compliance.} eLTER mandates FAIR data principles
    (Findable, Accessible, Interoperable, Reusable). The platform provides
    accessibility and interoperability through OGC STA, but lacks persistent identifiers
    (DOIs), structured provenance records, and machine-readable licensing metadata.
\end{itemize}

These gaps are architectural in nature and cannot be addressed by configuration changes
alone. They represent a substantial development effort that is beyond the scope of
this thesis but constitutes a realistic path for future work, given that the
foundational components (STA data access, OIDC authentication, containerised
deployment) are already in place.

\subsection{Sustainability of the TSM-TEMP Workarounds}
\label{sec:test-deployment-tsmtemp}

Section~\ref{sec:impl-tsm-assessment} assessed the TSM-TEMP workarounds from an
implementation perspective. The codebase contains 19~marked locations, concentrated
in \texttt{setup\_user\_database.py} and \texttt{sta\_views\_local/}. From a
sustainability perspective, two observations are relevant.

First, the local STA views have no operational disadvantages relative to the
SMS-backed alternative: they query tables the platform already manages, avoid
foreign-table lookups, and require no additional infrastructure. The migration path
back to upstream views remains documented, but the practical benefit of reverting is
questionable.

Second, the upstream SMS service has remained non-functional throughout the duration
of this thesis, and no public roadmap for its restoration has been communicated. This
reinforces the conclusion stated in Section~\ref{sec:impl-tsm-assessment}: a targeted
rewrite of the affected components --- replacing them with a clean implementation
rather than workarounds labeled as temporary --- is the more defensible long-term
strategy than waiting for an upstream fix with no guaranteed timeline.
