\chapter{Conclusion}
\label{chap:conclusion}

This thesis designed and implemented a platform for real-time environmental sensor data
collection, integration, and visualisation, building on the UFZ Time.IO solution as a
foundation and extending it to meet the operational requirements of Czech University of
Life Sciences Prague within the context of the European eLTER research infrastructure.

\medskip\noindent
The primary contributions of this work are:

\begin{enumerate}
  \item \textbf{TSM Orchestration Adaptation.}
        A working adaptation of the UFZ \texttt{tsm-orchestration} stack with local STA views
        implementing the missing FROST compatibility layer, modular Docker~Compose structure
        enabling overlay-based configuration, and Podman rootless support enabling deployment
        on security-constrained infrastructure.

  \item \textbf{The \texttt{water-dp} Platform.}
        A new application layer providing project-oriented sensor management, MQTT-driven
        alert evaluation, geospatial layer integration via GeoServer, and a multilingual web
        portal in English, Czech, and Slovak.

  \item \textbf{The \texttt{hydro-deploy} Deployment Layer.}
        A unified deployment orchestrator that merges the two stacks via Docker~Compose
        multi-file merge, manages secrets, validates cross-stack configuration consistency,
        and provides a scripted operator interface via a Makefile.
\end{enumerate}

The design decision that most significantly shaped the outcome was the choice to keep
\texttt{tsm-orchestration} and \texttt{water-dp} as separately deployable stacks rather than
merging them into a monolithic codebase. This separation preserved the ability to incorporate
upstream TSM updates while allowing the application layer to evolve independently. The
experience of this thesis, however, demonstrated that the upstream Time.IO project was not
responsive to the needs of external adopters: the SMS service remained non-functional, the
frontends produced invalid MQTT messages, and integration-layer documentation was absent. The
\texttt{water-dp} layer was built precisely to compensate for these deficiencies.

\section*{Limitations}
\addcontentsline{toc}{section}{Limitations}

The most significant limitation is the continued reliance on the \texttt{TSM-TEMP}
workarounds. The 19~marked locations in \texttt{tsm-orchestration} bypass the non-functional
upstream SMS service with local PostgreSQL views. While these views are operationally sound
(Section~\ref{sec:impl-tsm-assessment}), they carry the label of temporary fixes against a
codebase whose upstream shows no indication of addressing the root cause.

Test coverage of the \texttt{water-dp} backend, while comprehensive for the core service
layer (797~tests; Section~\ref{sec:test-approach-backend}), has gaps in the FROST client integration
(14\% coverage for \texttt{async\_frost\_client}) and the SMS service adapter (14\%). These
modules depend heavily on external services that are difficult to test without a running
FROST-Server and Keycloak instance.

Performance has not been benchmarked at scale beyond the CZU sensor fleet. The primary
bottleneck identified --- synchronous FROST-Server query processing
(Section~\ref{sec:test-perf-bottleneck}) --- would become critical at deployments with
thousands of Things. Quantitative throughput measurements under realistic multi-site load
remain future work.

Finally, the platform is not yet eLTER-ready. The architectural foundations are aligned
(OGC~STA data access, OIDC authentication, containerised deployment), but the
eLTER-specific requirements identified in Section~\ref{sec:test-deployment-elter} ---
metadata schemas, cross-site federation, vocabulary alignment, and FAIR compliance ---
represent substantial development effort beyond the scope of this thesis.

\section*{Future Work}
\addcontentsline{toc}{section}{Future Work}

The findings of this thesis point toward a clear direction for future development: rather
than continuing to patch the upstream Time.IO codebase through workarounds, the more
sustainable path is to take the \emph{architectural ideas} of Time.IO --- per-Thing schema
isolation, MQTT-driven worker pipelines, OGC~STA as the interoperability standard --- and
rewrite the problematic components from scratch, purpose-built for eLTER adoption.

The concrete future work items are:

\begin{enumerate}
  \item \textbf{TSM component rewrite.} Replace the \texttt{TSM-TEMP} workarounds
        and the non-functional SMS service with a clean STA view layer that is maintained
        as part of the platform rather than labelled as temporary. The rewrite scope is
        well-defined: 19~marked locations in one Python module and one SQL directory.

  \item \textbf{eLTER metadata integration.} Extend the project and sensor models
        with eLTER-specific metadata (DEIMS-SDR identifiers, EnvThes vocabulary bindings,
        habitat classifications) to enable the platform to serve as an eLTER site node.

  \item \textbf{Cross-site federation.} Implement a federation layer for aggregated
        queries across independently operated platform instances --- a prerequisite for
        eLTER's distributed, multi-site architecture.

  \item \textbf{FAIR data compliance.} Add persistent identifiers (DOIs), structured
        provenance records, and machine-readable licensing metadata to support the FAIR
        data principles mandated by eLTER.

  \item \textbf{Performance optimisation.} Introduce a caching layer (Redis) for
        frequently accessed FROST metadata, evaluate FROST-Server alternatives or
        clustering, and conduct load testing at scale beyond the CZU sensor fleet.

  \item \textbf{Test coverage completion.} Extend the automated test suite to cover the
        FROST client integration and SMS service adapter, potentially using
        containerised test fixtures (Testcontainers) for FROST-Server and Keycloak.
\end{enumerate}

\medskip\noindent
The OGC SensorThings API proved to be a well-chosen interoperability standard for this class
of environmental monitoring platform. Its entity model maps naturally onto the sensor data
management domain, and the availability of the FROST-Server implementation reduced the
initial integration effort. The primary friction was not with the standard itself but with
the state of the upstream implementation layer that was intended to bridge it to operational
sensor networks. The experience of this thesis suggests that the concept behind Time.IO ---
a modular, MQTT-driven, STA-compliant sensor data platform --- is sound, but a purpose-built
implementation targeting eLTER's specific requirements would be more effective than continued
adaptation of an upstream codebase that does not prioritise external adopters.
the per-thing schema model to FROST --- a gap that motivated the local STA views approach and that remains the
most important open item for future development.
